This file will be used to perform the same WRTDS and DLM analysis done to the dissolved oxygen concentration. This will also use Henry's Law 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(forecast)
library(EGRET)
library(EGRETci)
library(MARSS)
library(marssTMB)
library(tidyverse)
library(dplyr)
library(data.table)
library(smwrBase)
library(timetk)
library(miscTools)
library(RSocrata)
source('./functions/get_socrata_data_func.R')
```

This file will not bother assigning some of the data as it assumes all the oxygen code has been run and everything is already in the environment

```{r Temperature Plots}
# Do a quick plot of the data
GreenRiverDataExpanded %>%
  ggplot(aes(x=CollectDate, y=Temperature)) +
  geom_line() +
  ggtitle("Green River Temperature")

```


```{r Treat Data}
# Pull out Just the stuff we need and replace the 0's with empty values
GreenTemp <- subset(GreenRiverDataExpanded, select = c(CollectDate, Year, Month, Temperature))
GreenTemp <- replace(GreenTemp, GreenTemp==0.00, NA)
GreenTemp <- GreenTemp %>% 
  rename(Date = CollectDate,
         ConcAve = Temperature)

# Add all of the missing fields for WRTDS
GreenTemp$ConcLow <- GreenTemp$ConcAve
GreenTemp$ConcHigh <- GreenTemp$ConcAve
GreenTemp$DecYear <- decimal_date(GreenTemp$Date)
GreenTemp$SinDY <- sin(2*pi*GreenTemp$DecYear)
GreenTemp$CosDY <- cos(2*pi*GreenTemp$DecYear)
GreenTemp$Uncen <- NA
GreenTemp$Uncen <- c(1)
GreenTemp$Date <- as_date(GreenTemp$Date)
GreenTemp$RefDate <- as_date("1850-01-01")
GreenTemp$Julian<- difftime(GreenTemp$Date, GreenTemp$RefDate, units = "days")
GreenTemp$Julian <- as.integer(GreenTemp$Julian)
GreenTemp$MonthSeq <- interval(GreenTemp$RefDate, GreenTemp$Date) %/% months(1)
GreenTemp$waterYear <- waterYear(GreenTemp$Date, numeric = FALSE)

# Set the window to be analyized
yr_frst <- as_date("1980-01-01")
yr_last <- as_date("2022-12-31")
GreenTemp <- GreenTemp[GreenTemp[, "Date"] >= yr_frst & GreenTemp[, "Date"] <= yr_last,]
row.names(GreenTemp) <- NULL

#Quick plot of new data
GreenTemp %>%
  ggplot(aes(x=Date, y=ConcAve)) +
  geom_line() +
  ggtitle("Green River Temp")
```




The WRTDS/EGRET model is intended to be used on dissolved nutrients. It will be interesting to see how it handles temperature.
```{r WRTDS/EGRET model}
Sample2 <- na.omit(GreenTemp) # Create the sample data frame
row.names(Sample2) <- NULL

INFO2<- readNWISInfo(siteID,"00010")
INFO2$shortName <- "Green River at Aurburn, WA"


eList2 <- mergeReport(INFO2, Daily, Sample2) # Merge the data with the special EGRET function

# Run WRTDS/EGRET model:
eList2 <- modelEstimation(eList2)

# Test some of the plotting functions:
plotConcTimeDaily(eList2)
plotConcHist(eList2)
multiPlotDataOverview(eList = eList2)
errorStats(eList2) 
```
R-Squared Value is 0.835, the flux is 0.925








```{r }
# Merge the water quality (Sample) and the discharge (Daily) frames
# Discharge dataframe has daily records, while the quality frame has bi-monthly
# This means the response variable will have lots of missing entries
tempFrame <- data_frame()
tempFrame <- full_join(Sample2, Daily, by = "Date") 
tempFrame <- arrange(tempFrame,tempFrame$Date)
tempFrame$SinDY <- sin(2*pi*tempFrame$DecYear.y) 
tempFrame$CosDY <- cos(2*pi*tempFrame$DecYear.y) 

# Compress all the data into monthly averages
yearFrame2 <- tempFrame %>% mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(decyear = as.double((mean(DecYear.y))))
QFrame2 <- tempFrame %>% mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(LogQ = as.double(mean(LogQ))) 
sineFrame2 <- tempFrame %>%  mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(SinDY = as.double(mean(SinDY)))
cosFrame2 <- tempFrame %>%  mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(CosDY = as.double(mean(CosDY)))
TFrame<- tempFrame %>%  mutate(month = as.Date(Date, "%B")) %>%
  group_by(MonthSeq.y) %>%
    summarize(logTemp = as.double(log(mean(ConcAve, na.rm = TRUE))))

TT = nrow(TFrame)

# Response variable: log(Dissolved Oxygen)
TFrame$logTemp[is.nan(TFrame$logTemp)] <- NA
Temp = t(TFrame$logTemp)

#Covariates: Time(as the decimal year), Daily ave discharge (logQ), seasonal effects (sin and cos of the decimal year)
Z = array(NA,c(1,5,TT))
Z[1,1,] = c(1)
Z[1,2,] = yearFrame2$decyear
Z[1,3,] = QFrame2$LogQ
Z[1,4,] = sineFrame2$SinDY
Z[1,5,] = cosFrame2$CosDY

modlistT = list( #all other arguments are left at default
  Z = Z,
  U = "zero",
  A = "zero",
  B = "identity",
  Q = ldiag(list(5e-6,"q2","q3","q4","q5"))
)

# initial effects were set to 0, maybe 1 would be better
# EEH: This approach is not working; need to use a year effect; see below
marssfitT <- MARSS(Temp, model = modlistT, inits = list(x0 = matrix(c(0),5,1)),  control = list(maxit=3000)
                , method = "TMB"
                 )
```

```{r Residuals}
autoplot.marssMLE(marssfitT,silent = TRUE)
Z2=Z
dim(Z2)= c(5,dim(Z)[3])
A = Z2[,1:3] %*% marssfit1$states[1:3,]
plot(diag(A))

TT=dim(Z)[3]
lev = rep(NA, TT)
for (i in 1:TT){
  lev[i] = Z[1, 1:3,i] %*% marssfitT$states[1:3,i]
}
plot(lev)
apply(Z,2,function(x){any(is.na(x))})

```

```{r}
# x is a vector of 2 numbers
findpeak <- function(x){
  a <- x[1]; b <- x[2]
  if(a > 0 & b > 0) angle <- atan(a/b)
  if(a > 0 & b < 0) angle <- pi/2 + atan(a/-b)
  if(a < 0 & b < 0) angle <- pi + atan(-a/-b)
  if(a < 0 & b > 0) angle <- 3*pi/2 + atan(-a/b)
  if(a == 0 & b > 0) angle <- 0
  if(a == 0 & b < 0) angle <- pi
  if(b == 0 & a > 0) angle <- pi/2
  if(b == 0 & a < 0) angle <- 3*pi/2
  if(a == 0 & b == 0) angle <- NA
  return(12*angle/(2*pi))
}
```

So it looks like something happened and the peak shifted by a few days.
Scale is on y axis is months.

```{r}
peak <- apply(marssfitT$states[4:5,],2,findpeak)
plot(peak, type="l")
```


You can also look now at the amplitude: low to high. This one is easy.
Amplitude is $a^2 + b^2$. This one shows that the cycle amplitude has
generally stayed the same except whatever weird thing happened.

```{r}
amplitude <- sqrt(marssfitT$states[4,]^2 + marssfitT$states[5,]^2)
plot(amplitude, type="l")
```