This file will be used to perform the same WRTDS and DLM analysis done to the dissolved oxygen concentration. This will also use Henry's Law 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(forecast)
library(EGRET)
library(EGRETci)
library(MARSS)
library(marssTMB)
library(tidyverse)
library(dplyr)
library(data.table)
library(smwrBase)
library(timetk)
library(miscTools)
library(RSocrata)
source('./functions/get_socrata_data_func.R')
```

This file will not bother assigning some of the data as it assumes all the oxygen code has been run and everything is already in the environment

```{r Temperature Plots}
# Do a quick plot of the data
GreenRiverDataExpanded %>%
  ggplot(aes(x=CollectDate, y=Temperature)) +
  geom_line() +
  ggtitle("Green River Temperature")

```


```{r Treat Data}
# Pull out Just the stuff we need and replace the 0's with empty values
GreenTemp <- subset(GreenRiverDataExpanded, select = c(CollectDate, Year, Month, Temperature))
GreenTemp <- replace(GreenTemp, GreenTemp==0.00, NA)
GreenTemp <- GreenTemp %>% 
  rename(Date = CollectDate,
         ConcAve = Temperature)

# Add all of the missing fields for WRTDS
GreenTemp$ConcLow <- GreenTemp$ConcAve
GreenTemp$ConcHigh <- GreenTemp$ConcAve
GreenTemp$DecYear <- decimal_date(GreenTemp$Date)
GreenTemp$SinDY <- sin(2*pi*GreenTemp$DecYear)
GreenTemp$CosDY <- cos(2*pi*GreenTemp$DecYear)
GreenTemp$Uncen <- NA
GreenTemp$Uncen <- c(1)
GreenTemp$Date <- as_date(GreenTemp$Date)
GreenTemp$RefDate <- as_date("1850-01-01")
GreenTemp$Julian<- difftime(GreenTemp$Date, GreenTemp$RefDate, units = "days")
GreenTemp$Julian <- as.integer(GreenTemp$Julian)
GreenTemp$MonthSeq <- interval(GreenTemp$RefDate, GreenTemp$Date) %/% months(1)
GreenTemp$waterYear <- waterYear(GreenTemp$Date, numeric = FALSE)

# Set the window to be analyized
yr_frst <- as_date("1980-01-01")
yr_last <- as_date("2022-12-31")
GreenTemp <- GreenTemp[GreenTemp[, "Date"] >= yr_frst & GreenTemp[, "Date"] <= yr_last,]
row.names(GreenTemp) <- NULL

#Quick plot of new data
GreenTemp %>%
  ggplot(aes(x=Date, y=ConcAve)) +
  geom_line() +
  ggtitle("Green River Temp")
```




The WRTDS/EGRET model is intended to be used on dissolved nutrients. It will be interesting to see how it handles temperature.
```{r WRTDS/EGRET model}
Sample2 <- na.omit(GreenTemp) # Create the sample data frame
Sample2 <- distinct(Sample2)
row.names(Sample2) <- NULL

eList2 <- mergeReport(INFO, Daily, Sample2) # Merge the data with the special EGRET function

# Run WRTDS/EGRET model:
eList2 <- modelEstimation(eList2)

# Test some of the plotting functions:
plotConcTimeDaily(eList2)
plotConcHist(eList2)
multiPlotDataOverview(eList = eList)
errorStats(eList2) 
```









```{r }
# Merge the water quality (Sample) and the discharge (Daily) frames
# Discharge dataframe has daily records, while the quality frame has bi-monthly
# This means the response variable will have lots of missing entries
tempFrame <- data_frame()
tempFrame <- full_join(Sample, Daily, by = "Date") 
tempFrame <- arrange(tempFrame,tempFrame$Date)
tempFrame$SinDY <- sin(2*pi*tempFrame$DecYear.y) 
tempFrame$CosDY <- cos(2*pi*tempFrame$DecYear.y) 

# Compress all the data into monthly averages
yearFrame2 <- tempFrame %>% mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(decyear = as.double((mean(DecYear.y))))
QFrame2 <- tempFrame %>% mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(LogQ = as.double(mean(LogQ))) 
sineFrame2 <- tempFrame %>%  mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(SinDY = as.double(mean(SinDY)))
cosFrame2 <- tempFrame %>%  mutate(month = as.Date(Date, "%B")) %>%
    group_by(MonthSeq.y) %>%
    summarize(CosDY = as.double(mean(CosDY)))
TFrame<- tempFrame %>%  mutate(month = as.Date(Date, "%B")) %>%
  group_by(MonthSeq.y) %>%
    summarize(logTemp = as.double(log(mean(ConcAve, na.rm = TRUE))))

TT = nrow(TFrame)

# Response variable: log(Dissolved Oxygen)
TFrame$logTemp[is.nan(TFrame$logTemp)] <- NA
Temp = t(TFrame$logTemp)

#Covariates: Time(as the decimal year), Daily ave discharge (logQ), seasonal effects (sin and cos of the decimal year)
Z = array(NA,c(1,5,TT))
Z[1,1,] = c(1)
Z[1,2,] = yearFrame2$decyear
Z[1,3,] = QFrame2$LogQ
Z[1,4,] = sineFrame2$SinDY
Z[1,5,] = cosFrame2$CosDY

modlistT = list( #all other arguments are left at default
  Z = Z,
  U = "zero",
  A = "zero",
  B = "identity",
  Q = ldiag(list("q1","q2","q3","q4","q5"))
)

# initial effects were set to 0, maybe 1 would be better
# EEH: This approach is not working; need to use a year effect; see below
marssfitT <- MARSS(Temp, model = modlistT, inits = list(x0 = matrix(c(0),5,1)),  control = list(maxit=3000)
                , method = "TMB"
                 )
```

```{r Residuals}
autoplot.marssMLE(marssfitT,silent = TRUE)


```



# Ideal DO Estimation Using Henry's Law
To test the hypothesis that the DO is following fluxuations in temperature, I will use the temperature data to calculate estimates of DO according to Henry's Gas law. I will use the following equations, provided by D. Tromans (2000)
$$ln(k)=[0.046T^2 + 203.35T \times ln(T/298) - (299.378 + 0.092T)(T-298)-20.591 \times 10^3]/[8.3144\times T]$$
$$ C=k\times P(O2) \space | \space C(molO2/kgH2O),\space P(atm)$$
$$ DO(mg/L) = C \times 32,000 mg/mol \space O2 \times 0.99975 kg/L \space H2O $$


```{r Henry's Law}
# Start with the raw temperature data, We will assume the partial oxygen pressure remains constant at 0.21 atm to keep from overcomplicating things
Sample2$KelvinTemp <- Sample2$ConcAve + 273.15

Sample2$HenryK <- exp((0.046*(Sample2$KelvinTemp^2) + 203.35*Sample2$KelvinTemp*log(Sample2$KelvinTemp/298) - (299.378 + 0.092*Sample2$KelvinTemp)*(Sample2$KelvinTemp - 298) - 20.591*10^3)/(8.3144*Sample2$KelvinTemp))

Sample2$DOest <- Sample2$HenryK * 0.21 * 32000 * 0.99975

ggplot() +
  geom_point(data = Sample, aes(x = Date, y = ConcAve), color = "blue") + # must include argument label "data"
  geom_point(data = Sample2, aes(x = Date, y = DOest), color = "red") +
  ggtitle("Green River DO, Observed vs Henry's law Estimate")

```
A couple of things stick out of this plot:
1. There are a few instances where the water is apparently supersaturated with oxygen. This is interesting but I am not sure how important it is to what I need to get done right now
2. The maximum concentrations line up pretty closely until 2005 when observed maximum values become lower more consistently. I would like to ask some of the Fish folks if they can help me use some sort of method to determine if this behavior is statistically significant. There seems to be a change to the eye, but the eye is unreliable.
3. The minimum observed concentrations are consistently lower than the predicted concentrations. These are during the warmest times on the year. It could be attributable to a few things. The low flow could lead to the water not being mixed as well, and therefor hampering gas exchange. The warmer water could have a higher contaminant concentration in it and therefor a larger BOD and more respiration. There could be a higher concentration of ions and metals that reduce the solubility of oxygen

In-phase with Henry's law and the more out-of-phase as we continue?